---
title: "Accuracy vs RT" 
output: html_document
classoption: landscape
knit: (function(inputFile, encoding) { 
      out_dir <- 'FunctionResults2';
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_file=file.path(dirname(inputFile), out_dir, 'AccuracyVsRTAnalysis.html')) })
---

This script is really superceded by the scripts in "/Users/Olsonac-admin/Documents/current/PKU/pku_function_groups" which allow different groupings of functions to be compared.  We've kept this just for comparison/archive purposes.

```{r setup}
# this script runs a meta-analysis for a single function (e.g. 'Flexibility and planning')
# NOTE:  change the "title:" above to the name of the measure you are analyzing
#        There are some similar variables that need to be set below so labels are correct
#        see the comment #  SET THE TASK TITLE HERE # below
#        and the comment for the plot filename in the same region

# the first time you may need to install the library 'metafor'
# run the following line after removing the '#' then add the '#' back once you've installed the package
# install.packages("metafor")

# added heterogeneity measure and pooled effect size (was just z-stat and p-value)

#rm(list = ls()) # clear global environment

# load libraries 
library(metafor)
library(metaviz)
library(kableExtra)
library(tidyverse)
library(altmeta)
library(broom)

options(digits = 10) # make sure we have sufficient digits when writing .csv files 

# set the working directory to the folder with the data
# make this match the folder where you've put the files
CurDir<-"~/Documents/current/PKU/pku_cog_functions_meta_analysis/May2022_meta-analysis_with_added_papers"
knitr::opts_knit$set(root.dir=CurDir)



setwd(CurDir)
source("ForestPlotFunctions.R")
```

```{r}
RawData<-read.csv("DataForMetaAnalysis.csv")
head(RawData)
```

```{r}
AccRTStudies<-read.csv("Accuracy_vs_ReactionTime_StudyList.csv")
head(AccRTStudies)
```

```{r}
########################  SET THE TASK TITLE HERE ##################################
# label for this file
TaskTitle<-"AccVsRT"   
####################### SET THE OUTPUT FILENAME FOR THE PLOT HERE ##################
TaskFilename<-"AccVsRT"  # change the label here to match the data -- this will be used to label the plot file output.
```


```{r}
RawData$StudyNameFactor<-as.factor(RawData$StudyName)
RawData$FunctionFactor<-as.factor(RawData$Function)
RawData$TaskNameFactor<-as.factor(RawData$ShortTitle)
RawData$StudyFactor <- as.factor(RawData$StudyFactor)

AccRTStudies$FunctionFactor<-factor(as.character(AccRTStudies$Function),levels=levels(RawData$FunctionFactor))
AccRTStudies$StudyNameFactor<-factor(as.character(AccRTStudies$StudyName),levels=levels(RawData$StudyNameFactor))
AccRTStudies$TaskNameFactor<-factor(as.character(AccRTStudies$ShortTitle),levels=levels(RawData$TaskNameFactor))
AccRTStudies$AccRT<-as.factor(AccRTStudies$FunctionGroup)

# just keep columns used in join
AccRTStudies<-AccRTStudies %>% select(FunctionFactor,StudyNameFactor,TaskNameFactor,AccRT)

PKUData<-left_join(AccRTStudies,RawData,by=c("FunctionFactor",
                                          "StudyNameFactor",
                                          "TaskNameFactor"),keep=FALSE)
```


```{r}
MyData<-PKUData
# write.csv(MyData,"Acc_vs_rt_raw_data.csv",row.names=FALSE) # for checking

# substitute mean if CurrentPhe is missing
MyData$CurrentPhe[which(is.na(MyData$CurrentPhe))]<-mean(MyData$CurrentPhe,na.rm=TRUE)


# variables are in the form:
# [data frame] $ [column name]  e.g. PKUData$ExpN refers to the PKU datq frame
# and the "ExpN" column (this has the N for the number of patients in a study)

head(MyData)
```

```{r}
StudySummary <- MyData %>% select(StudyName, ShortTitle, Function, AccRT) %>% group_by(StudyName,ShortTitle,Function,AccRT) %>% summarise(N=n())
print(StudySummary)
```


```{r}
ggplot(MyData,aes(x=AccRT,y=g)) + geom_boxplot() + geom_jitter(color="grey", alpha=0.5)
```

Note that these results are unweighted

```{r}
UnWeightedGSummary <- MyData %>% group_by(AccRT) %>% summarise(mean_g = mean(g), sd_g = sd(g), n_g = n())
print(UnWeightedGSummary)
```
**Warning** unweighted results -- for cross-checking ballpark figures with weighted results
            weighted results should be reported

```{r}
HetResult<-metahet(MyData$g,MyData$Vg)
ISqr<-HetResult$I2
Q<-HetResult$Q
TauSqr<-HetResult$tau2.DL
```

```{r}
# this step does the statistical model for the meta-analysis
# g is the effect sizes
# Vg is the variance of the effect sizes
# random is what describes the relationship between numbers and studies
# 1 | StudyNum / ObsNum means that ObsNum (different for each measure) is nested
# within StudyNum.  "1 |" is needed because this is how random intercepts are
# described when specifying a mixed-effect model.  1 means intercept and what follows
# the | is what the intercept is for. 1 | StudyNum / ObsNum means a random intercept for
# ObsNum nested within StudyNum

RandomTermString <- "~ 1 | StudyFactor / TaskNameFactor "

MRTAcc<-rma.mv(g,
           Vg,
           random = formula(RandomTermString), 
           mods = ~ AccRT,
           test="z", 
           data=MyData, 
           method="REML")
summary(MRTAcc)
```


```{r}
MyData$CenteredPhe<-MyData$CurrentPhe-mean(MyData$CurrentPhe)
```

```{r}
AccData<-MyData[MyData$AccRT == "accuracy",]

AccHetResult<-metahet(AccData$g,AccData$Vg)
AccISqr<-AccHetResult$I2
AccQ<-AccHetResult$Q
AccTauSqr<-AccHetResult$tau2.DL

AccSummary <-  AccData %>% summarise(mean_g = mean(g), sd_g=sd(g), N_g=n())
print(AccSummary)
```

```{r}
# look at acc with Phe

MAccPhe<-rma.mv(g,
            Vg,
            random = formula(RandomTermString), 
            test="z", 
            mods= ~ CenteredPhe,
            data=AccData, 
            method="REML")
summary(MAccPhe)
```


```{r}
# look at acc w/o Phe

MAcc<-rma.mv(g,
             Vg,
             random = formula(RandomTermString), 
             test="z", 
             data=AccData, 
             method="REML")
summary(MAcc)
```

```{r}
WtMatrix<-weights(MAcc,type="matrix")
WtRowSum<-apply(WtMatrix,1,sum)
AccData$Weights<-WtRowSum/sum(WtRowSum)
```

```{r}

WtMean2 <- sum(AccData$g * AccData$Weights)
MyP2 <- ggplot(AccData,aes(x=g, y=seq(1,length(g)), size=Weights)) + geom_point() + geom_vline(xintercept=WtMean2,color="red")
MyP2 <- MyP2 + annotate(geom="label", x = -0.75,y = 2,label = sprintf("Overall ES = %.2f",WtMean2),color = "red")
MyP2 <- MyP2 + ggtitle("Accuracy effect sizes weighted by variance")
MyP2
```


```{r}
print(WtMean2)
```


```{r}
PlotTableRaw <- AccData %>% select(StudyName, ShortTitle, ExpMean, 
                                   ExpSD, ExpN, ControlMean, 
                                   ControlSD, ControlN, Weights)
PTable<-data.frame(
  StudyName=sprintf("%s",PlotTableRaw$StudyName),
  MeasureName=sprintf("%s",PlotTableRaw$ShortTitle),
  ExpMean=sprintf("%6.2f",PlotTableRaw$ExpMean),
  ExpSD=sprintf("%6.2f", PlotTableRaw$ExpSD),
  ExpN=sprintf("%3d",PlotTableRaw$ExpN),
  ControlMean=sprintf("%6.2f",PlotTableRaw$ControlMean),
  ControlSD=sprintf("%6.2f", PlotTableRaw$ControlSD),
  ControlN=sprintf("%3d",PlotTableRaw$ControlN),
  Weight=sprintf("%.1f",PlotTableRaw$Weights*100)
  )

MinES <- -3 
MaxES <- 2

# different parts of the forest plot
PlotPercentage<- .35  # percentage of space devoted to forest plot (vs text)
StudyNamePercent<-0.25 # percent of _text_ area for study name
MeanPercent<-0.085 # width of columns for means - % of total _text_ area
SDPercent<-0.06 # width of column for SD
NPercent<-0.04 # width of column for N
WtPercent<-0.04

LittleSep<-0.002 # separation between columns in a group (% of _text_ area)
BigSep<-0.004 # separation between columns in different groups (% of _text_ area)
TextMargin<-0.01 # space between text and plot (% of text area)

ColumnSep<-c(LittleSep,LittleSep,LittleSep,LittleSep,BigSep,LittleSep,LittleSep,LittleSep)
ColumnWidths<-c(MeanPercent,SDPercent,NPercent,MeanPercent,SDPercent,NPercent,WtPercent)
MeasureWidth<-(1-TextMargin)-(sum(ColumnSep)+sum(ColumnWidths)+StudyNamePercent)
if(MeasureWidth < 0){
  c("******WARNING***** space for Measure is too small ************")
}
ColumnWidths<-c(MeasureWidth,ColumnWidths)
ColumnAlign<-c(4,2,2,2,2,2,2,2) # first is left aligned, others centered
HeaderAlign<-c(1,1,1,1,1,1,1,1) # center headers

ColumnPositionPercentages<-InterleaveSepsAndWidths(ColumnSep,ColumnWidths,ColumnAlign)
ColumnPositionPercentages<-ColumnPositionPercentages+StudyNamePercent

HeaderPositionPercentages<-InterleaveSepsAndWidths(ColumnSep,ColumnWidths,HeaderAlign)
HeaderPositionPercentages<-HeaderPositionPercentages+StudyNamePercent

HeaderLine1<-c("Measure","Mean","SD","N","Mean","SD","N","% Wt")
HeaderLine2<-c("","","PKU","","","Control","")
AxisTitle<-"PKU worse        Effect size        PKU better"

HeteroString<-sprintf(" Heterogeneity: Q(%2d) = %3.2f, p<%0.2f, I-squared = %0.3f, Tau-squared = %0.3f",
                      MAcc$k-1, MAcc$QE, MAcc$QEp, AccISqr, AccTauSqr)

TextDF<-PTable
MyCex<-0.94
ForestPlotWithSummary("Accuracy", "ForestPlots2/Accuracy.tiff", MAcc, PlotPercentage, StudyNamePercent, TextDF, ColumnPositionPercentages, ColumnAlign, HeaderPositionPercentages, HeaderLine1, HeaderLine2, MyCex, AxisTitle, HeteroString, MinES=NULL, MaxES=NULL)

```


```{r}
# look at rt with Phe
RTData<-MyData[MyData$AccRT == "reaction time",]

RTHetResult<-metahet(RTData$g,RTData$Vg)
RTISqr<-RTHetResult$I2
RTQ<-RTHetResult$Q
RTTauSqr<-RTHetResult$tau2.DL


MRTPhe<-rma.mv(g,
             Vg,
             random = formula(RandomTermString), 
             test="z", 
             mods = ~ CenteredPhe,
             data=RTData, 
             method="REML")
summary(MRTPhe)
```

```{r}
rt_summary <- RTData %>% summarise(mean_g = mean(g))
print(rt_summary)
```


```{r}
# look at  rt w/o Phe

MRT<-rma.mv(g,
             Vg,
             random = formula(RandomTermString), 
             test="z", 
             data=RTData, 
             method="REML")
summary(MRT)
```

```{r}
WtMatrix<-weights(MRT,type="matrix")
WtRowSum<-apply(WtMatrix,1,sum)
RTData$Weights<-WtRowSum/sum(WtRowSum)

PlotTableRaw <- RTData %>% select(StudyName, ShortTitle, ExpMean, ExpSD, ExpN, ControlMean, ControlSD, ControlN, Weights)
PTable<-data.frame(
  StudyName=sprintf("%s",PlotTableRaw$StudyName),
  MeasureName=sprintf("%s",PlotTableRaw$ShortTitle),
  ExpMean=sprintf("%6.2f",PlotTableRaw$ExpMean),
  ExpSD=sprintf("%6.2f", PlotTableRaw$ExpSD),
  ExpN=sprintf("%3d",PlotTableRaw$ExpN),
  ControlMean=sprintf("%6.2f",PlotTableRaw$ControlMean),
  ControlSD=sprintf("%6.2f", PlotTableRaw$ControlSD),
  ControlN=sprintf("%3d",PlotTableRaw$ControlN),
  Weight=sprintf("%.1f",PlotTableRaw$Weights*100)
  )

MinES <- -3 
MaxES <- 2

# different parts of the forest plot
PlotPercentage<- .35  # percentage of space devoted to forest plot (vs text)
StudyNamePercent<-0.25 # percent of _text_ area for study name
MeanPercent<-0.085 # width of columns for means - % of total _text_ area
SDPercent<-0.06 # width of column for SD
NPercent<-0.04 # width of column for N
WtPercent<-0.04

LittleSep<-0.002 # separation between columns in a group (% of _text_ area)
BigSep<-0.004 # separation between columns in different groups (% of _text_ area)
TextMargin<-0.01 # space between text and plot (% of text area)

ColumnSep<-c(LittleSep,LittleSep,LittleSep,LittleSep,BigSep,LittleSep,LittleSep,LittleSep)
ColumnWidths<-c(MeanPercent,SDPercent,NPercent,MeanPercent,SDPercent,NPercent,WtPercent)
MeasureWidth<-(1-TextMargin)-(sum(ColumnSep)+sum(ColumnWidths)+StudyNamePercent)
if(MeasureWidth < 0){
  c("******WARNING***** space for Measure is too small ************")
}
ColumnWidths<-c(MeasureWidth,ColumnWidths)
ColumnAlign<-c(4,2,2,2,2,2,2,2) # first is left aligned, others centered
HeaderAlign<-c(1,1,1,1,1,1,1,1) # center headers

ColumnPositionPercentages<-InterleaveSepsAndWidths(ColumnSep,ColumnWidths,ColumnAlign)
ColumnPositionPercentages<-ColumnPositionPercentages+StudyNamePercent

HeaderPositionPercentages<-InterleaveSepsAndWidths(ColumnSep,ColumnWidths,HeaderAlign)
HeaderPositionPercentages<-HeaderPositionPercentages+StudyNamePercent

HeaderLine1<-c("Measure","Mean","SD","N","Mean","SD","N","% Wt")
HeaderLine2<-c("","","PKU","","","Control","")
AxisTitle<-"PKU worse        Effect size        PKU better"

HeteroString<-sprintf(" Heterogeneity: Q(%2d) = %3.2f, p<%0.2f, I-squared = %0.3f, Tau-squared = %0.3f", 
                      MRT$k-1, MRT$QE, MRT$QEp, RTISqr, RTTauSqr)

TextDF<-PTable
MyCex<-0.94
ForestPlotWithSummary("Reaction time", "ForestPlots2/reaction_time.tiff", MRT, PlotPercentage, StudyNamePercent, TextDF, ColumnPositionPercentages, ColumnAlign, HeaderPositionPercentages, HeaderLine1, HeaderLine2, MyCex, AxisTitle, HeteroString, MinES=NULL, MaxES=NULL)

```

```{r}
MyData$Predicted1<-fitted(MRTAcc) # general model with Acc/RT moderator, but no Phe
```

```{r}
predicted_summary <- MyData %>% group_by(AccRT) %>% summarise(mean_predicted = mean(Predicted1), n_predicted = n())
print(predicted_summary)
```


```{r}
RPlot<-ggplot(MyData,aes(x=CurrentPhe,y=g,group=AccRT,color=AccRT))+geom_point()+geom_point(data=MyData[MyData$AccRT == "accuracy",],aes(x=CurrentPhe,y=Predicted1),color="black",shape=20) +geom_point(data=MyData[MyData$AccRT == "reaction time",],aes(x=CurrentPhe,y=Predicted1),color="black",shape=21)
print(RPlot) # rt is open circles
```

This model doesn't take Phe into account

```{r}
# this step does the statistical model for the meta-analysis
# g is the effect sizes
# Vg is the variance of the effect sizes
# random is what describes the relationship between numbers and studies
# 1 | StudyNum / ObsNum means that ObsNum (different for each measure) is nested
# within StudyNum.  "1 |" is needed because this is how random intercepts are
# described when specifying a mixed-effect model.  1 means intercept and what follows
# the | is what the intercept is for. 1 | StudyNum / ObsNum means a random intercept for
# ObsNum nested within StudyNum


# model with main effect of Phe and main effect of rt/acc, but no interaction
MRTAccPhe<-rma.mv(g,
           Vg,
           random = formula(RandomTermString), 
           mods= ~ MyData$CenteredPhe + MyData$AccRT,
           test="z", 
           data=MyData, 
           method="REML")
summary(MRTAccPhe)
```

1st line after intercept is the first moderator -- i.e. centered Phe, 
2nd line after the intercept is the second moderator -- i.e. Accuracy vs RT

```{r}
MyData$Predicted2<-fitted(MRTAccPhe)
```

```{r}
RPlot<-ggplot(MyData,aes(x=CurrentPhe,y=g,group=AccRT,color=AccRT))+geom_point()+geom_point(data=MyData[MyData$AccRT == "accuracy",],aes(x=CurrentPhe,y=Predicted2),color="black",shape=20) +geom_point(data=MyData[MyData$AccRT == "reaction time",],aes(x=CurrentPhe,y=Predicted2),color="black",shape=21)
RPlot<-RPlot + xlim(600,1500)
print(RPlot)
```

```{r}
# model with Phe X acc/rt interaction

MRTAccPheInter<-rma.mv(g,
           Vg,
           random = formula(RandomTermString), 
           mods= ~ CenteredPhe*AccRT,
           test="z", 
           data=MyData, 
           method="REML")
summary(MRTAccPheInter)
```

```{r}
MyData$Predicted3<-fitted(MRTAccPheInter)
```

```{r}
WtMatrix<-weights(MRTAccPheInter,type="matrix")
WtRowSum<-apply(WtMatrix,1,sum)
MyData$Weights3<-WtRowSum/sum(WtRowSum)

RPlot<-ggplot(MyData,aes(x=CurrentPhe,y=g,group=AccRT,color=AccRT,size=Weights3))+geom_point(alpha=0.5)+geom_point(data=MyData[MyData$AccRT == "accuracy",],aes(x=CurrentPhe,y=Predicted3),color="black",shape=20,size=3) +geom_point(data=MyData[MyData$AccRT == "reaction time",],aes(x=CurrentPhe,y=Predicted3),color="black",shape=21,size=3)
RPlot<-RPlot + xlim(600,1500)
print(RPlot)
```

```{r}
AIC(MRTAcc,MRTAccPhe,MRTAccPheInter)
```

```{r}
MyData$Predicted<-fitted(MRTAcc)
```

```{r}
PredDiffs<-MyData %>% group_by(AccRT) %>% summarise(PredES=mean(Predicted), N=n())
print(PredDiffs)
```

predicted accuracy effect size and rt effect size to see difference

```{r}
PredDiffs$PredES[PredDiffs$AccRT == "reaction time"] - PredDiffs$PredES[PredDiffs$AccRT == "accuracy"]
```


